---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# simplefeaturesbr

The goal of simplefeaturesbr is to make [Simple Features](https://en.wikipedia.org/wiki/Simple_Features) tidy data frames of Brazilian states and municipalities easily available in R.

## Installation

You can install simplefeaturesbr with:  

``` r
remotes::install_github("BrazilianPublicData/simplefeaturesbr")
```

## Usage

simplefeaturesbr has a main dataset, `br_municipalities`, and some simple convenience functions to select from this dataset. You can access the data with: 

```{r}
library(simplefeaturesbr)
data("br_municipalities")
``` 

It looks like this: 

```{r}
head(br_municipalities)
```

If you are interested in a specific state, use `select_state()`: 
```{r}
select_state("SP")
``` 

If you'd like to get data for a specific municipality, use `select_municipality()`:
```{r}
select_municipality("Sao Paulo")
``` 

There are two other things to help you. If you can't remember which [IBGE](https://www.ibge.gov.br/en/home-eng.html) code (two-letter abbreviations of Brazilian states) you need, use `UF()` (copy of a simple function from [congressbr](https://github.com/BrazilianPublicData/congressbr)):

```{r}
UF()
```

## Why?
Simple features make _much_ smaller dataframes than spatial polygons. You can plot them easily with ggplot2's `geom_sf()`: 

```{r}
library(dplyr); library(ggplot2)

select_municipality("Rio de Janeiro") %>% ggplot() + geom_sf()
```

Nice.
